Development of a educational and sustainable robotic platform for children and adults

### 2013-12-23 doc/paper.tex
autocommited by makefile

\documentclass[11pt,a4paper]{article}
%\documentclass[conference]{IEEEtran}
\usepackage[english]{babel} 
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage{eurosym}
\DeclareUnicodeCharacter{20AC}{\euro}
\usepackage{csquotes}
\usepackage{graphicx}
\usepackage{rotating} 
\usepackage{array}
\usepackage{booktabs} 
\usepackage{longtable}
\usepackage{ifthen}
\usepackage{xcolor}
\usepackage{tabu}
\usepackage{colortbl}
\usepackage{calc}
\usepackage{pifont}
\usepackage{forloop}
\usepackage[nomessages]{fp}

@@

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/30_dcmotor.jpg}
  \caption{Picture of a typical DC motor (CC-BY-SA Ben Oswald)}
\end{figure}

In terms of cost effectiveness a solution with two wheels powered by a DC motor each performs well. A DC motor with integrated gearbox (see figure @TODO) is available for 5.5 USD, the plastic wheel costs 3.5 USD, a motor driver such as the L293 costs about 3.5 USD rendering a total cost of 21.5 USD per robot. In addition speed sensors for each of the wheels are needed since the motors do not have exactly the same speed. Without speed control the robot cannot drive straight forward.

@@

\subsection{IR Reciever}
Infrared Reciever are special photodiodes which react on infrared light with a base frequency of 38kHz. To ease the use of IR recievers with microcontrollers, the photodiode in the IR reciever provides a near binary square voltage waveform instead of rising voltages. The cost of such a reciever is about 0.3USD.

@@

\subsection{Ultrasonic Distance Sensor}
Ultrasonic distance sensors sense distance between the sensor and an object in the visible area of the sensor. In case of ultrasonic sensors the visible area is the area where the echo spreads. By counting the time between the signal and the reflection the distance can be calculated using the speed of sound (340.29 m / s at sea level).

>>>>

\documentclass[11pt,a4paper]{article}
%\documentclass[conference]{IEEEtran}
\usepackage[english]{babel} 
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[utf8]{inputenc}
\usepackage{eurosym}
\DeclareUnicodeCharacter{20AC}{\euro}
\usepackage{csquotes}
\usepackage{graphicx}
\usepackage{rotating} 
\usepackage{array}
\usepackage{booktabs} 
\usepackage{longtable}
\usepackage{ifthen}
\usepackage{xcolor}
\usepackage{tabu}
\usepackage{colortbl}
\usepackage{calc}
\usepackage{pifont}
\usepackage{forloop}
\usepackage[nomessages]{fp}
\usepackage{caption}
\usepackage{subcaption}

@@

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/30_ballcaster.jpg}
  \caption{Metal ball caster (CC-BY-SA Ben Oswald)}
\end{figure}

@@

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{images/30_dcmotor.jpg}
  \caption{Standard DC motor}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{images/30_gearboxmotor.jpg}
  \caption{Gearbox motor with wheel}
  \end{subfigure}
  \caption{Different DC motor types (CC-BY-SA Ben Oswald)}
\end{figure}

In terms of cost effectiveness a solution with two wheels powered by a DC motor each performs well. A DC motor with integrated gearbox is available for 5.5 USD, the plastic wheel costs 3.5 USD, a motor driver such as the L293 costs about 3.5 USD rendering a total cost of 21.5 USD per robot. In addition speed sensors for each of the wheels are needed since the motors do not have exactly the same speed since they are subject to variation during the production process. Without speed control sensors the robot may not drive straight forward. A simple sensor based on a rotating encoder disk and a light beam which is interrupted by the disk is available from 0.3 USD and requires one microcontroller pin per wheel (a microcontoller pin with interrupt support is the most reliable solution).

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/30_speedsensor.jpg}
  \caption{Speedsensor (CC-BY-SA Ben Oswald)}
\end{figure}

@@

\subsection{IR Reciever}
Infrared Reciever are special photodiodes which react on infrared light with a base frequency of 38kHz. To ease the use of IR recievers with microcontrollers, the photodiode in the IR reciever has an integrated Schmitt trigger and therefore provides a near binary square voltage waveform instead of rising voltages. The cost of such a reciever is about 0.3USD.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.5\textwidth]{images/30_ir.jpg}
  \caption{IR Reciever with integrated Schmitt trigger (CC-BY-SA Ben Oswald)}
\end{figure}

@@

\subsection{IR Proximity Sensor}
Infrared proximity sensors can be used in various ways. When measuring distances using proximity sensors one has to choose the right sensor type since the range varies from a few centimeters to a few meters. Another use of IR proximity sensors is to measure brightness. This usually works with sensors designed for very small distances. When the IR light is emitted it reflects better on bright surfaces than on dark surfaces which do not reflect light. A photodiode for infrared wavelength changes its electrical behaviour on the influence of the light reflected which can be measured using the microcontrollers analog-to-digital converter.

\begin{figure}[h!]
  \centering
  \begin{subfigure}{0.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{images/30_proximity_front.jpg}
  \caption{front-view}
  \end{subfigure}
  \begin{subfigure}{0.48\textwidth}
  \centering
  \includegraphics[width=0.9\linewidth]{images/30_proximity_back.jpg}
  \caption{side-view}
  \end{subfigure}
  \caption{IR proximity sensor (CC-BY-SA Ben Oswald)}
\end{figure}

Applications for proximity sensors are line following, edge detection (for instance of a table) or generic brightness measurement. When facing the front of the robot proximity sensors can be used to prevent collisions by measuring the proximity of the surrounding objects.

The price of a single proximity sensor is about 0.5 USD for average quality low distance sensors suitable for the application mentioned before. Microcontroller pin requirements are one pin to enable the infrared led within the proximity sensor (optional) and another analog-to-digital converter enabled pin to perform the readings of the photodiode. When multiple proximity sensors are used within a robot, analog multiplexer

\subsection{Ultrasonic Distance Sensor}
Ultrasonic distance sensors sense distance between the sensor and an object in the visible area of the sensor. In case of ultrasonic sensors the visible area is the area where the echo spreads. By counting the time between the signal and the reflection the distance can be calculated using the speed of sound (340.29m/s at sea level).
